---
title: "Debug STRING Enrichment Type Mismatch"
author: "Antigravity Debugger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(purrr)
library(tidyr)
```

## 1. Load Checkpointed Data

Replace the path below with the actual path to your generated `debug_enrichment_tables_raw.rds` file.

```{r load_data}
# Path to the checkpointed RDS file
# Example: "path/to/project/enrichment/string_db/debug_enrichment_tables_raw.rds"
debug_file_path <- "debug_enrichment_tables_raw.rds"

if (!file.exists(debug_file_path)) {
  stop("Checkpoint file not found! Please ensure you have run the analysis and the file was generated.")
}

enrichment_list <- readRDS(debug_file_path)
message("Successfully loaded enrichment list with ", length(enrichment_list), " contrasts.")
```

## 2. Inspect Column Types

Let's see which columns have conflicting types across different contrasts.

```{r inspect_types}
# Get all unique column names
all_cols <- unique(unlist(lapply(enrichment_list, colnames)))

# Create a summary table of types per column across contrasts
type_summary <- map_dfr(all_cols, function(col) {
  # Get classes for this column across all tables
  classes <- map_chr(enrichment_list, function(df) {
    if (col %in% colnames(df)) class(df[[col]])[1] else "Missing"
  })
  
  # Return row with column name and unique types found
  data.frame(
    column = col,
    unique_types = paste(unique(classes), collapse = ", "),
    has_conflict = length(unique(classes[classes != "Missing"])) > 1,
    stringsAsFactors = FALSE
  )
})

# Filter for conflicts
conflicts <- type_summary %>% filter(has_conflict)
knitr::kable(conflicts, caption = "Detected Column Type Conflicts")
```

## 3. Test Harmonization Logic

We will now apply the same dynamic harmonization used in the package to see if it resolves the conflicts safely.

```{r test_harmonization}
# Whitelist of columns we definitely want to be numeric
numeric_whitelist <- c(
  "enrichmentScore", "falseDiscoveryRate", "genesMapped", 
  "backgroundGenes", "genesInSet", "expected", 
  "observed", "pvalue", "fdr", "bonferroni", "proteinRanks"
)

harmonized_list <- enrichment_list

if (length(harmonized_list) > 1) {
  for (col in all_cols) {
    # Check types
    col_types <- sapply(harmonized_list, function(df) {
      if (col %in% colnames(df)) class(df[[col]])[1] else NA_character_
    })
    unique_types <- unique(na.omit(col_types))
    
    if (length(unique_types) > 1) {
      # Conflict detected! Resolve by coercion.
      target_type <- if (col %in% numeric_whitelist) "numeric" else "character"
      message(sprintf("Harmonizing column '%s' (%s) to %s.", 
                      col, paste(unique_types, collapse="/"), target_type))
      
      harmonized_list <- map(harmonized_list, function(df) {
        if (col %in% colnames(df)) {
          if (target_type == "numeric") {
            df[[col]] <- as.numeric(as.character(df[[col]]))
          } else {
            df[[col]] <- as.character(df[[col]])
          }
        }
        df
      })
    }
  }
}

# Verify success
tryCatch({
  merged_table <- bind_rows(harmonized_list, .id = "comparison")
  message("Success! bind_rows completed after harmonization.")
  head(merged_table)
}, error = function(e) {
  message("Failure! bind_rows still failing: ", e$message)
})
```

## 4. Manual Inspection of specific column

If it still fails, inspect the problematic column directly.

```{r manual_inspect}
# Change this to the column name that failed
problematic_col <- "proteinRanks" 

for (contrast_name in names(enrichment_list)) {
  df <- enrichment_list[[contrast_name]]
  if (problematic_col %in% colnames(df)) {
    cat(sprintf("\nContrast: %s\nType: %s\nValues (head):\n", 
                contrast_name, class(df[[problematic_col]])[1]))
    print(head(df[[problematic_col]]))
  }
}
```
